Embedding co-affiliation trajectories reveals structure of global scientific mobility

### 2020-01-30 paper/ic2s2-2020/ic2s2_mobility_emb.tex
Updated text of ic2s2 paper

Scientific mobility is closely linked to impact~\autocite{sugimoto_scientists_2017} and is an essential component of the scientific ecosystem.
This mobility is shaped by many factors, including geography, prestige, and policy\autocite{deville_career_2014,clauset_systematic_2015}.
Such complexity has resulted in a fragmented understanding of scientific mobility, limited to specific regions and mobility types.

\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{../Figs/ic2s2-2020/ic2s2_2020_mainfig.pdf}
	\caption{ 
	\textbf{Embedding offer robust model of co-affiliation, reveals multi-scale structure.}
	\textbf{a.} Examples of how ``sentences'' for the skip-gram model are created from author's affiliations listed on papers.
	\textbf{b.} Organization sentences are used as input for word embedding model. Each organization is represented as a vector.
	\textbf{c.} Organization embedding better captures closeness (flux) than does geographic distance. 
 $P_{i}$ and $P_{j}$ are the number of researchers in organization $i$ and $j$, and $F_{ij}$ is the symmetric flow between them.
	Black points are the mean flux for each binned by distance or similarity. 
	\textbf{d.} Cosine similarity better predictor of flux than geographic distance, using gravity law.
	Box-plots show distribution of actual flux binned across predicted, colored by the degree to which it overlaps $x = y$. 
	Cosine similarity is better predictor when considering only pairs of organizations within the same country (\textbf{e}) and only pairs in different countries (\textbf{f}). 
	\textbf{g.} UMAP visualization of embedding demonstrates geographic structure.
	Visualization of vector subsets reveal additional structure for the \textbf{h.} U.S. and \textbf{i.} Massachusetts. 
	 }
	\label{fig:image}
\end{figure}

We leverage recent advances in neural networks and representation learning to construct a dense, meaningful, and coherent vector-space embedding of global scientific mobility, using individual mobility trajectories documented in bibliographic data. 
This representation facilitates the holistic study of scientific mobility, capturing the complex relationships between organizations around the world.

@@

We visualize the embedding using UMAP~\autocite{mcinnes_umap_2018} and observe strong geographic clustering (Fig~\ref{fig:image}.g), while also reflecting a mix of historical, religious, and linguistic relationships between countries. 
For example, South American countries are near Spain and Portugal and organizations in French-speaking Quebec and North Africa are nearer France. 
By visualizing subsets of vectors, we reveal more granular structure.
For example, the structure of U.S. universities demonstrates state-level geographic clustering (Fig~\ref{fig:image}.h) roughly organized by state and region, whereas Massachusetts (Fig~\ref{fig:image}.i) instead shows structure based on urban centers (Boston vs. Worcester), organization type (hospitals tend to be adjacent) and different university systems (UMass system cluster separate from Harvard, MIT, etc.).

@@

%
% Bibliography
%
%\bibliographystyle{acm}
\begingroup
\setstretch{0.8}
\small
\setlength\bibitemsep{1pt}
{\renewcommand{\markboth}[2]{}% Remove header adjustment
\printbibliography}
\endgroup

>>>>

Scientific mobility is closely linked to impact, innovation, and knoweldge diffusion~\autocite{sugimoto_scientists_2017, wagner_new_2008}.
Central to mobility is the notion of proximity~\autocite{torre_proximity_2005} between places — in the case of scientific mobility, research organizations. 
However, geography itself in insufficient to understand mobility, and other forms of proximity are too intangible to meaningfully measure~\autocite{boschma_proximity_2005}, 
resulting in a fragmented and incomplete understanding of scientific mobility.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.97\textwidth]{../Figs/ic2s2-2020/ic2s2_2020_mainfig.pdf}
	\caption{ 
	\textbf{Embedding offer robust model of co-affiliation, reveals multi-scale structure.}
	\textbf{a.} Examples of how ``sentences'' for the skip-gram model are created from author's affiliations listed on papers.
	\textbf{b.} Organization sentences are used as input for word embedding model. Each organization is represented as a vector.
	\textbf{c.} Organization embedding better captures closeness (flux) than does geographic distance. 
 $P_{i}$ and $P_{j}$ are the number of researchers in organization $i$ and $j$, and $F_{ij}$ is the symmetric flow between them.
	Black points are the mean flux for each binned by distance or similarity. 
	\textbf{d.} Cosine similarity better predictor of flux than geographic distance, using gravity law.
	Box-plots show distribution of actual flux binned across predicted, colored by the degree to which it overlaps $x = y$. 
	Cosine similarity is better predictor when considering only pairs of organizations within the same country (\textbf{e}) and only pairs in different countries (\textbf{f}). 
	\textbf{g.} UMAP visualization of embedding demonstrates geographic structure.
	Visualization of vector subsets reveal additional structure for the \textbf{h.} U.S. and \textbf{i.} Massachusetts. 
	 }
	\label{fig:image}
\end{figure}

We leverage recent advances in neural networks and representation learning to construct a dense, meaningful, and coherent vector-space embedding of global scientific mobility, using individual mobility trajectories documented in bibliographic data. 
This representation implicitely captures organizational and cognitive proximity between organizations~\autocite{boschma_proximity_2005}, allowing for the holistic study of scientific mobility.

@@

We visualize the embedding using UMAP~\autocite{mcinnes_umap_2018} and observe strong geographic clustering (Fig~\ref{fig:image}.g), while also reflecting a mix of historic, cultural, and linguistic relationships between countries. 
For example, South American countries are near Spain and Portugal and organizations in French-speaking Quebec and North Africa are nearer France. 
By visualizing subsets of vectors, we reveal more granular structure.
For example, the structure of U.S. universities demonstrates state-level geographic clustering (Fig~\ref{fig:image}.h) roughly organized by state and region, whereas states such as Massachusetts (Fig~\ref{fig:image}.i), California, and New York (not shown) instead show structure based on urban centers (Boston vs. Worcester), organization type (hospitals tend to be adjacent) and different university systems (UMass system cluster separate from Harvard, MIT, etc.).

@@

%
% Bibliography
%
%\bibliographystyle{acm}
\begingroup
\setstretch{0.8}
\setlength\bibitemsep{1pt}
{
\small
\renewcommand{\markboth}[2]{}% Remove header adjustment
\printbibliography
}
\endgroup
